Algorithms for Planar Maximum Covering Location by Ellipses Problems\tnoteref{t1}

### 2020-03-22 defesa/article/conclusion.tex
wip

In this work, we studied two problems of maximum planar covering using ellipses, with most of the substantial results being for the less previously-studied problem where the ellipses can be freely rotated, which we referred to as MCER. 
We devoted most of our work for the development of an exact algorithm for this problem, which depended on the development of an algorithm for a never-studied-before geometric subproblem, which we referred to as E3P. The algorithm proposed for this last problem involves determining the eigenvalues of a six by six complex matrix to obtain the roots of a degree-six polynomial. We believe that further properties can be discovered which would allow the development of a simpler method to obtain the roots of the polynomial whose roots are solutions of E3P.

An exact algorithm was also developed for the other problem where the ellipses had to be axis-parallel. We took a different approach compared to the methods proposed by \cite{canbolat, andreta}, and based our algorithm entirely on the model proposed for the Euclidean PMCLP in \cite{church:1984}.

After giving improvement suggestions for the implementation of both algorithms, we analyzed several numerical experiments that were made using instances from \cite{canbolat, andreta}, and some new ones. Both algorithms performed really well, obtaining optimal solutions for every previously known instance, and only timing out for the really challenging ones from the set of newly constructed ones.

>>>>

In this work, we developed new algorithms for two planar maximum covering by ellipses problems.
After some improvements suggestions were given, numerical experiments were conducted showing the efficiency of both algorithms for instances from previous works and for new ones with larger demand and facility sets.

We devoted most of our work to the development of an exact algorithm for the version of the problem where the ellipses can be freely rotated.
In the midst of that, a never-studied-before subproblem came up, for which we also proposed an algorithm involving determining the roots of a complex polynomial.
Our algorithm for this version of the problem was able to obtain optimal solutions for instances which the heuristic method developed in \cite{andreta} returned non-optimal ones, and the deterministic method timed out.

We believe that the never-studied-before subproblem, which was the subject of \autoref{section:e3p}, has some properties that we were not able to explore, which could be the subject of future work. 
We observed that the solutions, except for one case, always come in pairs; and that the generated complex polynomial whose roots provide solutions for the problem appears to be a self-inverse polynomial.

### 2020-03-22 defesa/article/implementation.tex
wip

In the algorithm to obtain every solution of E3P described in \autoref{section:e3p}, we assumed that a procedure which returns every eigenvalue of a given complex matrix was available. In practice, we used the famous linear algebra package LAPACK (see \cite{lapack} for more details).
Even though LAPACK is a library for the FORTRAN programming language, its routines can be made available in a C/C++ environment by simply adding the -llapack linking flag to the compilation. The only remarks, though, are that FORTRAN represents matrices in a column-major fashion, and receives parameters only by reference. Therefore, matrices must be transposed before being passed to a routine, and every parameter must receive a pointer to a variable containing its value.

@@

The actual coefficients of that polynomial would take more than ten pages if they were to be presented here, because of that, we made them available elsewhere in \url{https://sites.icmc.usp.br/andretta/tedeschi-2020/}.

>>>>

In the algorithm to obtain every solution of E3P described in \autoref{section:e3p}, we assumed that a procedure which returns every eigenvalue of a given complex matrix was available. In practice, we used the famous linear algebra package LAPACK (see \cite{lapack} for more details).
Even though LAPACK is a library for the FORTRAN programming language, its routines can be made available in a C/C++ environment by simply adding the \textit{-llapack} linking flag to the compilation. The only remarks, though, are that FORTRAN represents matrices in a column-major fashion, and receives parameters only by reference. Therefore, matrices must be transposed before being passed to a routine, and every parameter must receive a pointer to a variable containing its value.

@@

%The actual coefficients of that polynomial would take more than ten pages if they were to be presented here, because of that, we made them available elsewhere in \url{https://sites.icmc.usp.br/andretta/tedeschi-2020/}.

### 2020-03-22 defesa/article/introduction.tex
wip

It is fair to say that PMCLP with elliptical coverage has not been vastly studied as only two articles have been found on it. In \cite{canbolat}, a mixed-integer non-linear programming method was proposed as a first approach to the problem. For some instances, the method took too long and did not find an optimal solution. Because of that, a heuristic method was developed using a technique called Simulated Annealing, which was able to obtain solutions for the instances proposed in that study.
The problem was further explored in \cite{andreta}, which introduced the version where the ellipses can be freely rotated, to which an exact and a heuristic method was proposed, and developed a new method for the axis-parallel version of the problem, which was able to obtain optimal solutions for instances that the method proposed by \cite{canbolat} could not.
The exact method for the version with rotation could not obtain optimal solutions within a predefined time limit for several instances, the heuristic method though returned solutions for every instance, and impressively enough, obtained optimal solutions for every verifiable instance.

>>>>

It is fair to say that PMCLP with elliptical coverage has not been vastly studied as only two articles have been found on it. In \cite{canbolat}, a mixed-integer non-linear programming method was proposed as a first approach to the problem, and, because it took too long and did not find an optimal solution for some of the proposed instances, a heuristic method was developed using a technique called Simulated Annealing. 
%For some instances, the method took too long and did not find an optimal solution. Because of that, a heuristic method was developed using a technique called Simulated Annealing, which was able to obtain solutions for the instances proposed in that study.

Initially, in the version of the problem defined in \cite{canbolat}, the ellipses could not be rotated. In \cite{andreta} though, the problem where the ellipses can be freely rotated was introduced.

which introduced the version where the ellipses can be freely rotated.

to which an exact and a heuristic method was proposed, and developed a new method for the axis-parallel version of the problem, which was able to obtain optimal solutions for instances that the method proposed by \cite{canbolat} could not.
The exact method for the version with rotation could not obtain optimal solutions within a predefined time limit for several instances, the heuristic method though returned solutions for every instance, and impressively enough, obtained optimal solutions for every verifiable instance.

### 2020-03-22 defesa/article/mce.tex
wip

\begin{prp}\label{lema:mce-mwc}
	Let $(\Pp, \Ww, \{(a, b)\})$ be an instance of MCE, and $q\in \R^2$. If $\Pp \cap E(q) = A$, then $q \in \cap_{p\in A} E(p)$ and for any $q' \in \cap_{p\in A} E(p)$, we have $A \subset E(q')$.
\end{prp}
\begin{proof}
	Let $u, v \in \R^2$, then we have that $u \in E(v)$ if, and only if $v \in E(u)$, which is proved simply by observing that:
	\begin{equation}\label{eq:peq}
	u \in E(v) \Leftrightarrow ||u-v||_{a, b, 0} \le 1 \Leftrightarrow v \in E(u).
	\end{equation}
	Therefore, if $p\in A$, then by \autoref{eq:peq}, $q \in E(p)$, which implies $q \in \cap_{p\in A} E(p)$.
	The second part of the lemma also follows from \autoref{eq:peq} as if $q' \in \cap_{p\in A} E(p)$, then $||q' - p||_{a,b,0} \le 1$ for all $p \in A$.
\end{proof}

By \autoref{lema:mce-mwc}, we can ignore solutions that are not in $\cap_{p\in A} E(p)$, for some $A \subset \Pp$. Because ellipses can be described by strictly convex norms, we will show that the set $\cap_{p\in A} E(p)$ has some convenient properties, which can be used to narrow down the solutions' search space even more.

@@

To prove that this is true, we resort to the results of \cite{bi}, which develops an algorithm to determine the intersection region of $n$ strictly convex disks of fixed radius. There, it is stated that this intersection region is bounded by the arcs of some of the disks that are part of the intersection, and also that the vertices of this region form a subset of the set of pairwise intersections of the $n$ given circles.
Following this result, we define the next lemma.

>>>>

\begin{prp}\label{lema:mce-mwc}
	Let $(\Pp, \Ww, \{(a, b)\})$ be an instance of MCE, and $q\in \R^2$. If $\Pp \cap E(q) = A$, then $q \in \cap_{p\in A} E(p)$ and for any $q' \in \cap_{p\in A} E(p)$, we have $A \subset E(q')$.
\end{prp}
\begin{proof}
	Let $u, v \in \R^2$, then we have the following equivalence:
	\begin{equation}\label{eq:peq}
	u \in E(v) \Leftrightarrow ||u-v||_{a, b, 0} \le 1 \Leftrightarrow v \in E(u).
	\end{equation}
	Therefore, if $p\in A$, by \autoref{eq:peq}, then $q \in E(p)$, which implies $q \in \cap_{p\in A} E(p)$.
	The second part of the lemma also follows from \autoref{eq:peq}. If $q' \in \cap_{p\in A} E(p)$, then $||q' - p||_{a,b,0} \le 1$ for all $p \in A$.
\end{proof}

By \autoref{lema:mce-mwc}, given a solution $q\in\R^2$, such that for all $A \subset \Pp$, $|A|\neq \emptyset$, $q \not \in \cap_{p\in A} E(p)$, then $\Pp \cap E(q) = \emptyset$. Therefore, we can focus on studying the intersection set $\cap_{p\in A} E(p)$, as solutions that are not in one do not cover any demand points.

@@

To prove that this is true for ellipses, we resort to the results of \cite{bi}, which develops an algorithm to determine the intersection region of $n$ strictly convex disks of fixed radius. There, it is stated that this intersection region is bounded by the arcs of some of the disks that are part of the intersection, and also that the vertices of this region form a subset of the set of pairwise intersections of the $n$ given circles.
Following this result, we define the next lemma.
